---
title: "Interaction Accumulation Curves"
author: "Elena Quintero"
date: "`r Sys.Date()`"
output: github_document
---

Here we present four examples for the computation of Interactions Accumulation Curves (IAC) present in Figure 2 of the manuscript.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, message=F, warning=F, include=F}
library(tidyverse)
library(fossil)
library(vegan)
library(knitr)
suppressPackageStartupMessages(library(tidylog))
library(summarytools)
```

### A. FOCAL OBSERVATIONS

Example of the frugivory assemblage of *Cecropia glaziovii* individual plants at Intervales Park (Sao Paulo, Brasil). Data obtained during frugivory course (Pedro Jordano, Marco Aurelio Pizo, Mauro Galetti).

```{r}
f.obs_raw <- read.csv ("data/Cecropia_data.csv", header = T, sep=";",)
```

Create variable named 'species' by pasting genera and species; and remove undetermined species

```{r}
f.obs <- f.obs_raw %>%
  dplyr::mutate(species = paste0(gen,".", sp.)) %>%
  dplyr::filter(gen!= "Indet")
```

Obtain observation matrix where each row represents one individual of *Cecropia*, and frugivores are presented in columns.

```{r}
cecropia.obs <- as.data.frame(t(f.obs[,7:34])) %>% `colnames<-`(f.obs$species)
```

**Sampling effort:** There were 28 individual plants that were observed for continuous 5 hours on the same day.

```{r}
#Generate sampling effort list
cecropia.samp.eff= rep(5,28)
```

Obtain Species Richness estimators: Chao, first order Jack knife, second order Jack knife and Bootstrap with their respective SE.

```{r}
kable(specpool(cecropia.obs))
```

where "Species" indicates the number of species encountered (in this case 37 frugivore species), and "n" indicates the number of sampling effort units (in this case 12 *Cecropia* individuals).

IAC curve:

```{r IAC_focal_observations}
curve_f.obs <- specaccum(cecropia.obs, method="random", permutations=100, w=cecropia.samp.eff, effort=5)

plot(curve_f.obs, ci.type="polygon", xvar = "effort",lwd=2, ci.lty=0,
     ci.col= rgb(116/255,169/255,207/255,.6), ylab= "Interaction richness", 
     xlab="Observation time (h/plant)", main="Focal Observations")

```

### B. CAMERA TRAPS

Example of frugivory assemblage at three *Juniperus phoenicia* forest stands in Do침ana Natural Reserve (Huelva, Spain). Data from Jorge Isla's PhD Thesis.

```{r}
cam_raw <- read.csv("data/Juniperus_cams.csv", header = T, sep=";")
```

Filter data by exclusive "feeding" behavior, and summarize data by area, fortnight and species detected.
```{r, message=F}
cam <- cam_raw %>%
  dplyr::filter(behavior=="FEEDING") %>%
  dplyr::group_by(area, fortnight, sp, ) %>%
  dplyr::summarise(n_interactions=n())
```

Convert to observation matrix where each row represent one fortnight of data collection in each one of the areas. Columns indicate frugivore species detected.

```{r convert to observation matrix sab}
sab_obs<-cam %>%
  tidyr::pivot_wider(names_from= sp, values_from= n_interactions, 
                     values_fill= list(count= 0)) %>%
  replace(is.na(.), 0)
```

**Sampling effort:** Cameras traps (n=X?) were revised every 15 days in each area
```{r}
#Generate sampling effort list
juniperus_samp_eff <- rep(15,83)
```

Obtain Species Richness estimators:
```{r general curve sab}
kable(specpool(sab_obs[,-c(1:2)],sab_obs$area))
```

where "Species" indicates the number of species encountered (in this case 14, 10 and 11 frugivore species for each *Juniperus* stand), and "n" indicates the number of sampling effort units (28, 28 and 27 fortnights in each *Juniperus* stand).

IAC curve:
```{r IAC_camera_traps}
curve_cam <- specaccum(sab_obs[,-c(1:2)], method="random", permutations=100, effort=15, w=juniperus_samp_eff)

plot(curve_cam, ci.type="polygon",col="#555555", lwd=2, ci.lty=0, 
     ci.col= rgb(116/255,169/255,207/255,.6), ylab="Interaction richness", 
     xvar = "effort",
     xlab="Cameras working time in days", main="Camera Traps")
```

### C. AVIAN MIST-NETTING

Example of frugivory assemblage at Hato Rat칩n in Do침ana National Park (Huelva, Spain) from diet samples obstained from mist-netted birds between 1980-1983. Data from Pedro Jordano's PhD Thesis.

```{r}
mist.net_raw <- read.csv("data/Interactions_HR_1980.csv", sep=";")
```

Give a unique ID number for each bird capture, calculate total number of seeds per bird species and give a unique name for each sp-sp pairwise interaction.
```{r}
mist.net <- mist.net_raw %>%
            dplyr::mutate(sample=row_number()) %>% 
            tidyr::pivot_longer(cols=4:20,names_to = "disperser", 
                                values_to= "count") %>% 
            dplyr::mutate(interaction= paste(SPECIES,disperser))
```

Convert to observation matrix where each row represents a bird captured (*i.e.*, a diet sample), and columns represent avian frugivores.

```{r}
mist.net <- mist.net %>% 
           tidyr::pivot_wider(id_cols=sample, names_from = interaction, 
                              values_from= count, 
                              values_fill= list(count= 0), 
                              names_repair= "unique")
```

**Sampling effort:** there was a total of 1808 birds captured of which 374 did not have fruits/seed in their diet.

Obtain Interactions richness estimators:
```{r}
kable(specpool(mist.net[,-1]))
```

where "Species" indicates the number of **interactions** encountered (in this case 125 unique frugivore-plant pariwise interactions), and "n" indicates the number of sampling effort units (in this case 1808 diet samples).

IAC curve:
```{r IAC_avian_mist_netting}
curve_mist.net <- specaccum(mist.net[,-1], method="random",permutations=100)

plot(curve_mist.net, ci.type="polygon",col="#555555", lwd=2, ci.lty=0, 
      ci.col= rgb(116/255,169/255,207/255,.6), ylab="Interaction richness", 
      xlab="Number of birds captured", main="Avian mist-netting")
```

### D. DNA-Barcoding

Example for frugivory assemblage of *Pistacia lentiscus* individual plants in La Mancha del Rabicano, El Puntal in Do침ana National Park (Huelva, Spain). Data from Elena Quintero's PhD Thesis.

```{r}
dna.bc_raw <- read.csv("data/Pistacia_barcoding_winter.csv", header = T, sep=",",) 
```

Convert to observation matrix where each row represent a individual sample analyzed (*i.e.*, fecal sample or regurgitated seed), and columns represent avian frugivores.
```{r}
dna.bc <- dna.bc_raw %>%
   dplyr::select(-X) %>%
   tidyr::pivot_wider(names_from = disperser, values_from = priority, 
                      values_fill= list(priority= 0))
```

Obtain Species Richness Estimators
```{r}
kable(specpool(dna.bc[,-c(1,2)]))
```

where "Species" indicates the number of species encountered (in this case 16 frugivore species), and "n" indicates the number of sampling effort units (in this case 1072 fecal/seed samples collected in seed traps settled under *Pistacia lentiscus* individuals).

IAC curve:
```{r IAC_DNA_barcoding}
curve_pun <- specaccum(dna.bc[,-c(1,2)], method="random",permutations=100)

plot(curve_pun, ci.type="polygon",col="#555555", lwd=2, ci.lty=0, 
     ci.col=rgb(116/255,169/255,207/255,.6), ylab="Interaction richness", 
     xlab="Number of samples analyzed", main="DNA-barcoding")
```

------------------------------------------------------------------------

```{r session, echo= FALSE}
#-------------------------------------------------------------------------------
# In Rmd: 
# Last update: 

cat("Last update: ", format(Sys.time(), "%a %b %d %X %Y"), "\n", "\n")

# Session
sessionInfo()
#-------------------------------------------------------------------------------

```
