---
title: "CIs Modularity-Hato Ratón"
author: "Pedro Jordano"
date: "11/18/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Matrices

Final matrix Names:

`mn= Original mist-netting matrix from HR`
`obs= Original censuses matrix from HR`
`HRcualmerging= Cualitative matrix sum`
`HRcuanmerging= Cuantitative matrix sum`
`HRgtmerging = Grand Total Standarization`
`HRminmaxmerging = Min-max scaled matrix`

 
```{r}
# Load matrices
getwd()
hr.mn<- read.csv("../Case_studies_datasets/hr_mn_clean.csv",header=T,sep=",",dec=".")
hr.obs<- read.csv("../Case_studies_datasets/hr_obs_clean.csv",header=T,sep=",",dec=".")

# Convert all interactions in numeric
mn <- hr.mn%>%column_to_rownames(var="AASPECIES")
#as.data.frame(lapply(hr.mn, as.numeric))

obs <- hr.obs%>%column_to_rownames(var="AASPECIES")
#as.data.frame(lapply((pl.cam),hr.obs, as.numeric))

```

# Combination methods

```{r}
#Qualitative combination
HRcuanmerging <- mn+obs
HRcualmerging= HRcuanmerging
HRcualmerging[HRcuanmerging!=0]= 1

#Quantitative sum
HRcuanmerging <- mn+obs

# Grand total
mn.gts<- mn/sum(mn)
obs.gts<- obs/sum(obs)
HRgtmerging<- (mn.gts+obs.gts)/2
sum(HRgtmerging)

# Min-Max scaling
mn.mms<- (mn-min(mn))/(max(mn)-min(mn))
obs.mms<- (obs-min(obs))/(max(obs)-min(obs))
HRminmaxmerging<- (mn.mms+obs.mms)/2

```


## Bootstrap estimates of network modularity (M)

We estimate the confidence intervals by resampling the five adjacency matrices
obtained from the different methods and computing their average, SE, and CIs for modularity.

We use function `m_boot` for this.

## Functions to load
### Function to estimate M from bootstrapped matrices.

```{r m_boot.function}
# Function for Bootstrap loop to estimate M for resampled matrices. ------------
# USE WITH: m_boot(mymat, myresamp). myresamp defaults to 99.
m_boot <- function (mymat, myresamp= 99)  {
	TIME <- Sys.time()
	# Resampling plan for bootstrapping n=99 adjacency matrices.
	m.boot<- NULL
	# mat is the input matrix for which M is tested
	# mlike is the observed mean M value (in general, mat@likelihood)
	# The loop. PUT THIS IN A BOOT LOOP 
	for (i in 1:myresamp) {
		mat1 <- sample_frac(mymat, 0.8, replace= T)
		mnmetrics <- computeModules(mat1, method="Beckett", 
								   deep= FALSE, deleteOriginalFiles= FALSE, 
								   steps= 1000, tolerance= 1e-10, 
								   experimental= FALSE, forceLPA= FALSE)
		m.boot<- rbind(m.boot, mnmetrics@likelihood)   
	}
	colnames(m.boot)<- c("MBoot") # m.boot saves the M values for bootstrapped 
								  # matrices.
	return(m.boot)
	#
	Sys.time() - TIME
	#
} 

```


# Modularity

Final matrix Names:

`mn= Original mist-netting matrix from HR`
`obs= Original censuses matrix from HR`
`HRcualmerging= Cualitative matrix sum`
`HRcuanmerging= Cuantitative matrix sum`
`HRgtmerging = Grand Total Standarization`
`HRminmaxmerging = Min-max scaled matrix`

### Empirical, observed matrices
#### Mist-netting

```{r M_mn}
M_mn <- computeModules(mn, method="Beckett", deep= FALSE, 
					   deleteOriginalFiles= FALSE, steps= 1000, 
					   tolerance= 1e-10, experimental= FALSE, forceLPA= FALSE)
M_mn@likelihood
# Msig(mn, M_bc@likelihood)

# m_boot(mn, 99)
# CHECK: m_boot is not saving object m.boot, so I'm resolving here... :(( Chapuza

# Outputs bootstrapped parameter differences values as dataframe. -------------
MObs<- M_mn@likelihood

HR_MN_Mboot <- as.data.frame(cbind(MObs, # colMeans(m.boot)))
									 colMeans(m.boot<-m_boot(mn, 99))))
rnd.se <- apply(m.boot,2,sd)/sqrt(length(m.boot))   # SE randomized parameters
HR_MN_Mboot<- cbind(HR_MN_Mboot, rnd.se)
ci1 <-  MObs-(1.962*HR_MN_Mboot$rnd.se)
ci2 <-  MObs+(1.962*HR_MN_Mboot$rnd.se)
# ci1 <-  apply(m.boot,2,mean)-(1.962*HR_MN_Mboot$rnd.se)
# ci2 <-  apply(m.boot,2,mean)+(1.962*HR_MN_Mboot$rnd.se)

HR_MN_Mboot<- cbind(HR_MN_Mboot, ci1, ci2)
colnames(HR_MN_Mboot)<- c("MObs", "meanM.boot", "SE", 
								  "CI_low",
								  "CI_high")
#
# Output parameters
# meandf         # Bootstrap mean of parameter values
# SE             # Bootstrap SE for parameter values
# CI_low         # Bootstrap CI (lower) 
# CI_high        # Bootstrap CI (higher) 
HR_MN_Mboot

```


#### Direct observations (spot-census along transects)

```{r M_obs}
M_CENSUS <- computeModules(obs, method="Beckett", deep= FALSE, 
					   deleteOriginalFiles= FALSE, steps= 1000, 
					   tolerance= 1e-10, experimental= FALSE, forceLPA= FALSE)
M_CENSUS@likelihood
# Msig(obs, M_cam@likelihood)
# m_boot(obs, 99)
# CHECK: m_boot is not saving object m.boot, so I'm resolving here... :(( Chapuza

# Outputs bootstrapped parameter differences values as dataframe. -------------
MObs<- M_CENSUS@likelihood

HR_CENSUS_Mboot <- as.data.frame(cbind(MObs, # colMeans(m.boot)))
									 colMeans(m.boot<-m_boot(obs, 99))))
rnd.se <- apply(m.boot,2,sd)/sqrt(length(m.boot))   # SE randomized parameters
HR_CENSUS_Mboot<- cbind(HR_CENSUS_Mboot, rnd.se)
ci1 <-  MObs-(1.962*HR_CENSUS_Mboot$rnd.se)
ci2 <-  MObs+(1.962*HR_CENSUS_Mboot$rnd.se)
# ci1 <-  apply(m.boot,2,mean)-(1.962*HR_CENSUS_Mboot$rnd.se)
# ci2 <-  apply(m.boot,2,mean)+(1.962*HR_CENSUS_Mboot$rnd.se)

HR_CENSUS_Mboot<- cbind(HR_CENSUS_Mboot, ci1, ci2)
colnames(HR_CENSUS_Mboot)<- c("MObs", "meanM.boot", "SE", 
								  "CI_low",
								  "CI_high")
#
# Output parameters
# meandf         # Bootstrap mean of parameter values
# SE             # Bootstrap SE for parameter values
# CI_low         # Bootstrap CI (lower) 
# CI_high        # Bootstrap CI (higher) 
HR_CENSUS_Mboot

```

### Merged matrices
#### Cualitative matrix sum

```{r M_HRcualmerging, warning=FALSE}
M_HRcualmerging <- computeModules(HRcualmerging, method="Beckett", 
								deep= FALSE, deleteOriginalFiles= FALSE, 
								steps= 1000, tolerance= 1e-10, 
								experimental= FALSE, forceLPA= FALSE)
M_HRcualmerging@likelihood
# Msig(HRcualmerging, M_HRcualmerging@likelihood)

# m_boot(HRcualmerging, 99)
# CHECK: m_boot is not saving object m.boot, so I'm resolving here... :(( Chapuza

# Outputs bootstrapped parameter differences values as dataframe. -------------
MObs<- M_HRcualmerging@likelihood

HRcualmerging_Mboot <- as.data.frame(cbind(MObs, # colMeans(m.boot)))
									 colMeans(m.boot<-m_boot(HRcualmerging, 99))))
rnd.se <- apply(m.boot,2,sd)/sqrt(length(m.boot))   # SE randomized parameters
HRcualmerging_Mboot<- cbind(HRcualmerging_Mboot, rnd.se)
ci1 <-  MObs-(1.962*HRcualmerging_Mboot$rnd.se)
ci2 <-  MObs+(1.962*HRcualmerging_Mboot$rnd.se)
# ci1 <-  apply(m.boot,2,mean)-(1.962*HRcualmerging_Mboot$rnd.se)
# ci2 <-  apply(m.boot,2,mean)+(1.962*HRcualmerging_Mboot$rnd.se)

HRcualmerging_Mboot<- cbind(HRcualmerging_Mboot, ci1, ci2)
colnames(HRcualmerging_Mboot)<- c("MObs", "meanM.boot", "SE", 
								  "CI_low",
								  "CI_high")
#
# Output parameters
# meandf         # Bootstrap mean of parameter values
# SE             # Bootstrap SE for parameter values
# CI_low         # Bootstrap CI (lower) 
# CI_high        # Bootstrap CI (higher) 
HRcualmerging_Mboot

```


#### Cuantitative matrix sum

```{r M_HRcuanmerging, warning=FALSE}
M_HRcuanmerging <- computeModules(HRcuanmerging, method="Beckett", 
								deep= FALSE, deleteOriginalFiles= FALSE, 
								steps= 1000, tolerance= 1e-10, 
								experimental= FALSE, forceLPA= FALSE)
M_HRcuanmerging@likelihood
# Msig(HRcuanmerging, M_HRcuanlmerging@likelihood)

# m_boot(HRcuanmerging, 99)
# CHECK: m_boot is not saving object m.boot, so I'm resolving here... :(( Chapuza

# Outputs bootstrapped parameter differences values as dataframe. -------------
MObs<- M_HRcuanmerging@likelihood
HRcuanmerging_Mboot <- as.data.frame(cbind(MObs, # colMeans(m.boot)))
									 colMeans(m.boot <- m_boot(HRcuanmerging, 99))))
rnd.se <- apply(m.boot,2,sd)/sqrt(length(m.boot))   # SE randomized parameters
HRcuanmerging_Mboot<- cbind(HRcuanmerging_Mboot, rnd.se)
ci1 <-  MObs-(1.962*HRcuanmerging_Mboot$rnd.se)
ci2 <-  MObs+(1.962*HRcuanmerging_Mboot$rnd.se)
# ci1 <-  apply(m.boot,2,mean)-(1.962*HRcuanmerging_Mboot$rnd.se)
# ci2 <-  apply(m.boot,2,mean)+(1.962*HRcuanmerging_Mboot$rnd.se)

HRcuanmerging_Mboot <- cbind(HRcuanmerging_Mboot, ci1, ci2)
colnames(HRcuanmerging_Mboot)<- c("MObs", "meanM.boot", "SE", 
								  "CI_low",
								  "CI_high")
#
# Output parameters
# meandf         # Bootstrap mean of parameter values
# SE             # Bootstrap SE for parameter values
# CI_low         # Bootstrap CI (lower) 
# CI_high        # Bootstrap CI (higher) 
HRcuanmerging_Mboot

```


#### Grand Total Standarization

```{r M_HRgtmerging, warning=FALSE}
M_HRgtmerging <- computeModules(HRgtmerging, method="Beckett", 
								deep= FALSE, deleteOriginalFiles= FALSE, 
								steps= 1000, tolerance= 1e-10, 
								experimental= FALSE, forceLPA= FALSE)
M_HRgtmerging@likelihood
# Msig(HRgtmerging, M_HRcualmerging@likelihood)

# m_boot(HRgtmerging, 99)
# CHECK: m_boot is not saving object m.boot, so I'm resolving here... :(( Chapuza

# Outputs bootstrapped parameter differences values as dataframe. -------------
MObs<- M_HRgtmerging@likelihood
HRgtmerging_Mboot <- as.data.frame(cbind(MObs, # colMeans(m.boot)))
								   colMeans(m.boot<- m_boot(HRgtmerging, 99))))
rnd.se <- apply(m.boot,2,sd)/sqrt(length(m.boot))   # SE randomized parameters
HRgtmerging_Mboot<- cbind(HRgtmerging_Mboot, rnd.se)
ci1 <-  MObs-(1.962*HRgtmerging_Mboot$rnd.se)
ci2 <-  MObs+(1.962*HRgtmerging_Mboot$rnd.se)
# ci1 <-  apply(m.boot,2,mean)-(1.962*HRgtmerging_Mboot$rnd.se)
# ci2 <-  apply(m.boot,2,mean)+(1.962*HRgtmerging_Mboot$rnd.se)


HRgtmerging_Mboot<- cbind(HRgtmerging_Mboot, ci1, ci2)
colnames(HRgtmerging_Mboot)<- c("MObs", "meanM.boot", "SE", 
								   "CI_low",
								   "CI_high")
#
# Output parameters
# meandf         # Bootstrap mean of parameter values
# SE             # Bootstrap SE for parameter values
# CI_low         # Bootstrap CI (lower) 
# CI_high        # Bootstrap CI (higher) 
HRgtmerging_Mboot

```


#### Min-max scaled matrix

```{r M_HRminmaxmerging, warning=FALSE}
M_HRminmaxmerging <- computeModules(HRminmaxmerging, method="Beckett", 
								deep= FALSE, deleteOriginalFiles= FALSE, 
								steps= 1000, tolerance= 1e-10, 
								experimental= FALSE, forceLPA= FALSE)
M_HRminmaxmerging@likelihood
# Msig(HRminmaxmerging, M_HRcualmerging@likelihood)

# m_boot(HRminmaxmerging, 99)
# CHECK: m_boot is not saving object m.boot, so I'm resolving here... :(( Chapuza

# Outputs bootstrapped parameter differences values as dataframe. -------------
MObs<- M_HRminmaxmerging@likelihood
HRminmaxmerging_Mboot <- as.data.frame(cbind(MObs, # colMeans(m.boot)))
									   colMeans(m.boot<-m_boot(HRminmaxmerging, 99))))
rnd.se <- apply(m.boot,2,sd)/sqrt(length(m.boot))   # SE randomized parameters
HRminmaxmerging_Mboot<- cbind(HRminmaxmerging_Mboot, rnd.se)
ci1 <-  MObs-(1.962*HRminmaxmerging_Mboot$rnd.se)
ci2 <-  MObs+(1.962*HRminmaxmerging_Mboot$rnd.se)
# ci1 <-  apply(m.boot,2,mean)-(1.962*HRminmaxmerging_Mboot$rnd.se)
# ci2 <-  apply(m.boot,2,mean)+(1.962*HRminmaxmerging_Mboot$rnd.se)

HRminmaxmerging_Mboot<- cbind(HRminmaxmerging_Mboot, ci1, ci2)
colnames(HRminmaxmerging_Mboot)<- c("MObs", "meanM.boot", "SE", 
								   "CI_low",
								   "CI_high")
#
# Output parameters
# meandf         # Bootstrap mean of parameter values
# SE             # Bootstrap SE for parameter values
# CI_low         # Bootstrap CI (lower) 
# CI_high        # Bootstrap CI (higher) 
HRminmaxmerging_Mboot

```


------------------------------------------------------------------------------

```{r session, echo=T}
sessionInfo()

```



