---
title: "CIs-Modularity_El Puntal"
author: "Pedro Jordano"
date: "11/18/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Matrices

Final matrix Names:

`bc = Original Barcoding matrix from Pistacia lentiscus`   
`cualmerging = Cualitative matrix sum`        
`cuanmerging = Cuantitative matrix sum`          
`gtmerging = Grand Total Standarization`         
`minmaxmerging = Min-max scaled matrix`            
`seffortmerging = Sampling effort merging by mean among standardized matrices.`       


```{r load-data}
# Load matrices

pl.bc<- read.csv("../Case_studies_datasets/pl_bc_clean.csv",header=T,sep=",",dec=".")
pl.cam<- read.csv("../Case_studies_datasets/pl_cam_clean.csv",header=T,sep=",",dec=".")

# Convert all interactions in numeric
bc <- as.data.frame(lapply(pl.bc, as.numeric))%>% 
	column_to_rownames(var="X")
# ??column_to_rownames
cam <- as.data.frame(lapply(pl.cam, as.numeric)) %>% 
	        column_to_rownames(var="X")

# Qualitative combination
PUNcuanmerging<- bc+cam
PUNcualmerging= PUNcuanmerging
PUNcualmerging[PUNcuanmerging!= 0]= 1

# Quantitative sum
PUNcuanmerging<- bc+cam

# Grand total
bct<- bc/sum(bc)
camt<- cam/sum(cam)
PUNgtmerging<- (bct+camt)/2

# Min-Max scaling
bct<- (bc-min(bc))/(max(bc)-min(bc))
camt<- (cam-min(cam))/(max(cam)-min(cam))
PUNminmaxmerging<- (bct+camt)/2

# Sampling effort standardization
bceff<- read.csv("../../Data/pistacia/bc_sampling_effort.csv",header=T,sep=";",dec=".", comment.char = "#")
cameff<- read.csv("../../Data/pistacia/cam_sampling_effort.csv",header=T, sep=",",dec=".")

# I need to convert all data in visits/min/plant
# First TIME TRANSFORMATION
bct1 <- bc[,,]/(12*60*bceff[,7]) #I  HAVE NUMBER OF DAYS THE SEED TRAY WAS IN THE PLANT - DIVIDE BY DAYS*12*60 =renders visits/min (12 hours of the day, not 24)
camt1 <- cam[,,]/cameff[,3] # I HAVE MINS THE CAMS WAS RECODING

# Then AREA TRANSFORMATION 
bceff$correc_fac <- bceff$plant_area/bceff$area_seed_trap # I CALCULATE A 
                             # CORRECTION FACTOR i.e. plant area/seed tray area
bct2 <- bct1[,,]*bceff[,8] #I multiply it by the feces found
camt2 <- camt1[,,]*2 # I could multiply CAM DATA by 2 since the camera has only 
                     # recorded half of the plant - SHOULD I? (Not included)
# Comment PJ: Not necessary to x2 given that we're not trying to estimate an
# actual visitation rate (i.e., a real value for visits), but just to standardize.
# By the way, a x2 would simply rescale all the values and that's not relevant 
# for our purpose here.

# Then merging
PUNseffortmerging<- (bct2+camt2)/2

```


## Bootstrap estimates of network modularity (M)

We estimate the confidence intervals by resampling the five adjacency matrices
obtained from the different methods and computing their average, SE, and CIs for modularity.

We use function `m_boot` for this.

Recall the final matrix Names:

`bc = Original Barcoding matrix from Pistacia lentiscus`            
`cam = Original GoPro matrix from Pistacia lentiscus`           
`cualmerging = Cualitative matrix sum`            
`cuanmerging = Cuantitative matrix sum`           
`gtmerging = Grand Total Standarization`            
`minmaxmerging = Min-max scaled matrix`           
`seffortmerging = Sampling effort merging by mean among standardized matrices.`        

## Functions to load
### Function to estimate M from bootstrapped matrices.

```{r m_boot.function}
# Function for Bootstrap loop to estimate M for resampled matrices. ------------
# USE WITH: m_boot(mymat, myresamp). myresamp defaults to 99.
m_boot <- function (mymat, myresamp= 99)  {
	TIME <- Sys.time()
	# Resampling plan for bootstrapping n=99 adjacency matrices.
	m.boot<- NULL
	# mymat is the input matrix for which M is tested
	# mlike is the observed mean M value (in general, mat@likelihood)
	# The loop. PUT THIS IN A BOOT LOOP 
	for (i in 1:myresamp) {
		mat1 <- sample_frac(mymat, 0.8, replace= T)
		mnmetrics <- computeModules(mat1, method="Beckett", 
								   deep= FALSE, deleteOriginalFiles= FALSE, 
								   steps= 1000, tolerance= 1e-10, 
								   experimental= FALSE, forceLPA= FALSE)
		m.boot<- rbind(m.boot, mnmetrics@likelihood)   
	}
	colnames(m.boot)<- c("MBoot") # m.boot saves the M values for bootstrapped 
								  # matrices.
	return(m.boot)
	#
	Sys.time() - TIME
	#
} 

```


# Modularity

### Empirical, observed matrices
#### DNA bar-coding

```{r M_bc}
M_bc <- computeModules(bc, method="Beckett", deep= FALSE, 
					   deleteOriginalFiles= FALSE, steps= 1000, 
					   tolerance= 1e-10, experimental= FALSE, forceLPA= FALSE)
M_bc@likelihood
# Msig(bc, M_bc@likelihood)

#m_boot(bc, 99)
# CHECK: m_boot is not saving object m.boot, so I'm resolving here... :(( Chapuza

# Outputs bootstrapped parameter differences values as dataframe. -------------
MObs<- M_bc@likelihood
PUN_BC_Mboot <- as.data.frame(cbind(MObs, 
									   colMeans(m.boot<-m_boot(bc, 99))))
rnd.se <- apply(m.boot,2,sd)/sqrt(length(m.boot))   # SE randomized parameters
PUN_BC_Mboot<- cbind(PUN_BC_Mboot, rnd.se)
ci1 <-  MObs-(1.962*PUN_BC_Mboot$rnd.se)
ci2 <-  MObs+(1.962*PUN_BC_Mboot$rnd.se)

PUN_BC_Mboot<- cbind(PUN_BC_Mboot, ci1, ci2)
colnames(PUN_BC_Mboot)<- c("MObs", "meanM.boot", "SE", 
								   "CI_low",
								   "CI_high")
#
# Output parameters
# meandf         # Bootstrap mean of parameter values
# SE             # Bootstrap SE for parameter values
# CI_low         # Bootstrap CI (lower) 
# CI_high        # Bootstrap CI (higher) 
PUN_BC_Mboot

```


#### Camera-traps

```{r M_cam}
M_cam <- computeModules(cam, method="Beckett", deep= FALSE, 
					   deleteOriginalFiles= FALSE, steps= 1000, 
					   tolerance= 1e-10, experimental= FALSE, forceLPA= FALSE)
M_cam@likelihood
# Msig(cam, M_cam@likelihood)
#m_boot(bc, 99)
# CHECK: m_boot is not saving object m.boot, so I'm resolving here... :(( Chapuza

# Outputs bootstrapped parameter differences values as dataframe. -------------
MObs<- M_cam@likelihood
PUN_CAM_Mboot <- as.data.frame(cbind(MObs, 
									   colMeans(m.boot<-m_boot(cam, 99))))
rnd.se <- apply(m.boot,2,sd)/sqrt(length(m.boot))   # SE randomized parameters
PUN_CAM_Mboot<- cbind(PUN_CAM_Mboot, rnd.se)
ci1 <-  MObs-(1.962*PUN_CAM_Mboot$rnd.se)
ci2 <-  MObs+(1.962*PUN_CAM_Mboot$rnd.se)

PUN_CAM_Mboot<- cbind(PUN_CAM_Mboot, ci1, ci2)
colnames(PUN_CAM_Mboot)<- c("MObs", "meanM.boot", "SE", 
								   "CI_low",
								   "CI_high")
#
# Output parameters
# meandf         # Bootstrap mean of parameter values
# SE             # Bootstrap SE for parameter values
# CI_low         # Bootstrap CI (lower) 
# CI_high        # Bootstrap CI (higher) 
PUN_CAM_Mboot

```

### Merged matrices
#### Cualitative matrix sum

```{r M_cualmerging, warning=FALSE}
M_cualmerging <- computeModules(PUNcualmerging, method="Beckett", 
								deep= FALSE, deleteOriginalFiles= FALSE, 
								steps= 1000, tolerance= 1e-10, 
								experimental= FALSE, forceLPA= FALSE)
M_cualmerging@likelihood
# Msig(PUNcualmerging, M_cualmerging@likelihood)

#m_boot(PUNcualmerging, 99)
# CHECK: m_boot is not saving object m.boot, so I'm resolving here... :(( Chapuza

# Outputs bootstrapped parameter differences values as dataframe. -------------
MObs<- M_cualmerging@likelihood
PUN_cualmerging_Mboot <- as.data.frame(cbind(MObs, 
									   colMeans(m.boot<-m_boot(PUNcualmerging, 99))))
rnd.se <- apply(m.boot,2,sd)/sqrt(length(m.boot))   # SE randomized parameters
PUN_cualmerging_Mboot<- cbind(PUN_cualmerging_Mboot, rnd.se)
ci1 <-  MObs-(1.962*PUN_cualmerging_Mboot$rnd.se)
ci2 <-  MObs+(1.962*PUN_cualmerging_Mboot$rnd.se)

PUN_cualmerging_Mboot<- cbind(PUN_cualmerging_Mboot, ci1, ci2)
colnames(PUN_cualmerging_Mboot)<- c("MObs", "meanM.boot", "SE", 
								   "CI_low",
								   "CI_high")
#
# Output parameters
# meandf         # Bootstrap mean of parameter values
# SE             # Bootstrap SE for parameter values
# CI_low         # Bootstrap CI (lower) 
# CI_high        # Bootstrap CI (higher) 
PUN_cualmerging_Mboot

```


#### Cuantitative matrix sum

```{r M_cuanmerging, warning=FALSE}
M_cuanmerging <- computeModules(PUNcuanmerging, method="Beckett", 
								deep= FALSE, deleteOriginalFiles= FALSE, 
								steps= 1000, tolerance= 1e-10, 
								experimental= FALSE, forceLPA= FALSE)
M_cuanmerging@likelihood
# Msig(PUNcuanmerging, M_cuanmerging@likelihood)

# m_boot(PUNcuanmerging, 99)
# CHECK: m_boot is not saving object m.boot, so I'm resolving here... :(( Chapuza

# Outputs bootstrapped parameter differences values as dataframe. -------------
MObs<- M_cuanmerging@likelihood
PUN_cuanmerging_Mboot <- as.data.frame(cbind(MObs, 
									   colMeans(m.boot<-m_boot(PUNcuanmerging, 99))))
rnd.se <- apply(m.boot,2,sd)/sqrt(length(m.boot))   # SE randomized parameters
PUN_cuanmerging_Mboot<- cbind(PUN_cuanmerging_Mboot, rnd.se)
ci1 <-  MObs-(1.962*PUN_cuanmerging_Mboot$rnd.se)
ci2 <-  MObs+(1.962*PUN_cuanmerging_Mboot$rnd.se)

PUN_cuanmerging_Mboot<- cbind(PUN_cuanmerging_Mboot, ci1, ci2)
colnames(PUN_cuanmerging_Mboot)<- c("MObs", "meanM.boot", "SE", 
								   "CI_low",
								   "CI_high")
#
# Output parameters
# meandf         # Bootstrap mean of parameter values
# SE             # Bootstrap SE for parameter values
# CI_low         # Bootstrap CI (lower) 
# CI_high        # Bootstrap CI (higher) 
PUN_cuanmerging_Mboot

```


#### Grand Total Standarization

```{r M_gtmerging, warning=FALSE}
M_gtmerging <- computeModules(PUNgtmerging, method="Beckett", 
								deep= FALSE, deleteOriginalFiles= FALSE, 
								steps= 1000, tolerance= 1e-10, 
								experimental= FALSE, forceLPA= FALSE)
M_gtmerging@likelihood
# Msig(PUNgtmerging, M_gtmerging@likelihood)

# m_boot(PUNgtmerging, 99)
# CHECK: m_boot is not saving object m.boot, so I'm resolving here... :(( Chapuza

# Outputs bootstrapped parameter differences values as dataframe. -------------
MObs<- M_gtmerging@likelihood
PUN_gtmerging_Mboot <- as.data.frame(cbind(MObs, 
									 colMeans(m.boot<-m_boot(PUNgtmerging, 99))))
rnd.se <- apply(m.boot,2,sd)/sqrt(length(m.boot))   # SE randomized parameters
PUN_gtmerging_Mboot<- cbind(PUN_gtmerging_Mboot, rnd.se)
ci1 <-  MObs-(1.962*PUN_gtmerging_Mboot$rnd.se)
ci2 <-  MObs+(1.962*PUN_gtmerging_Mboot$rnd.se)

PUN_gtmerging_Mboot<- cbind(PUN_gtmerging_Mboot, ci1, ci2)
colnames(PUN_gtmerging_Mboot)<- c("MObs", "meanM.boot", "SE", 
								   "CI_low",
								   "CI_high")
#
# Output parameters
# meandf         # Bootstrap mean of parameter values
# SE             # Bootstrap SE for parameter values
# CI_low         # Bootstrap CI (lower) 
# CI_high        # Bootstrap CI (higher) 
PUN_gtmerging_Mboot

```


#### Min-max scaled matrix

```{r M_minmaxmerging, warning=FALSE}
M_minmaxmerging <- computeModules(PUNminmaxmerging, method="Beckett", 
								deep= FALSE, deleteOriginalFiles= FALSE, 
								steps= 1000, tolerance= 1e-10, 
								experimental= FALSE, forceLPA= FALSE)
M_minmaxmerging@likelihood
# Msig(PUNminmaxmerging, M_minmaxmerging@likelihood)

# m_boot(PUNminmaxmerging, 99)
# CHECK: m_boot is not saving object m.boot, so I'm resolving here... :(( Chapuza

# Outputs bootstrapped parameter differences values as dataframe. -------------
MObs<- M_minmaxmerging@likelihood
PUN_minmaxmerging_Mboot <- as.data.frame(cbind(MObs, 
										 colMeans(m.boot<-m_boot(PUNminmaxmerging, 99))))
rnd.se <- apply(m.boot,2,sd)/sqrt(length(m.boot))   # SE randomized parameters
PUN_minmaxmerging_Mboot<- cbind(PUN_minmaxmerging_Mboot, rnd.se)
ci1 <-  MObs-(1.962*PUN_minmaxmerging_Mboot$rnd.se)
ci2 <-  MObs+(1.962*PUN_minmaxmerging_Mboot$rnd.se)

PUN_minmaxmerging_Mboot<- cbind(PUN_minmaxmerging_Mboot, ci1, ci2)
colnames(PUN_minmaxmerging_Mboot)<- c("MObs", "meanM.boot", "SE", 
								   "CI_low",
								   "CI_high")
#
# Output parameters
# meandf         # Bootstrap mean of parameter values
# SE             # Bootstrap SE for parameter values
# CI_low         # Bootstrap CI (lower) 
# CI_high        # Bootstrap CI (higher) 
PUN_minmaxmerging_Mboot

```


#### Sampling effort merging by mean among standardized matrices

```{r M_seffortmerging, warning=FALSE}
M_seffortmerging <- computeModules(PUNseffortmerging, method="Beckett", 
								deep= FALSE, deleteOriginalFiles= FALSE, 
								steps= 1000, tolerance= 1e-10, 
								experimental= FALSE, forceLPA= FALSE)
M_seffortmerging@likelihood
# Msig(PUNseffortmerging, M_seffortmerging@likelihood)

# m_boot(PUNseffortmerging, 99)
# CHECK: m_boot is not saving object m.boot, so I'm resolving here... :(( Chapuza

# Outputs bootstrapped parameter differences values as dataframe. -------------
MObs<- M_seffortmerging@likelihood
PUN_seffortmerging_Mboot <- as.data.frame(cbind(MObs, 
										  colMeans(m.boot<-m_boot(PUNseffortmerging, 99))))
rnd.se <- apply(m.boot,2,sd)/sqrt(length(m.boot))   # SE randomized parameters
PUN_seffortmerging_Mboot<- cbind(PUN_seffortmerging_Mboot, rnd.se)
ci1 <-  MObs-(1.962*PUN_seffortmerging_Mboot$rnd.se)
ci2 <-  MObs+(1.962*PUN_seffortmerging_Mboot$rnd.se)

PUN_seffortmerging_Mboot<- cbind(PUN_seffortmerging_Mboot, ci1, ci2)
colnames(PUN_seffortmerging_Mboot)<- c("MObs", "meanM.boot", "SE", 
								   "CI_low",
								   "CI_high")
#
# Output parameters
# meandf         # Bootstrap mean of parameter values
# SE             # Bootstrap SE for parameter values
# CI_low         # Bootstrap CI (lower) 
# CI_high        # Bootstrap CI (higher) 
PUN_seffortmerging_Mboot

```

------------------------------------------------------------------------------

```{r session, echo=T}
sessionInfo()

```



