---
title: "Network indices confidence intervals - Hato Ratón"
author: "Pedro Jordano"
date: "11/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Matrices

Final matrix Names:

`mn = Original mist-netting matrix from HR`

`obs = Original censuses matrix from HR`

`HRcualmerging = Cualitative matrix sum`

`HRcuanmerging = Cuantitative matrix sum`

`HRgtmerging = Grand Total Standarization`

`HRminmaxmerging = Min-max scaled matrix`


```{r}
# Load matrices
getwd()
hr.mn<- read.csv("../Case_studies_datasets/hr_mn_clean.csv",header=T,sep=",",dec=".")
hr.obs<- read.csv("../Case_studies_datasets/hr_obs_clean.csv",header=T,sep=",",dec=".")

# Convert all interactions in numeric
mn <- hr.mn%>%column_to_rownames(var="AASPECIES")
#as.data.frame(lapply(hr.mn, as.numeric))

obs <- hr.obs%>%column_to_rownames(var="AASPECIES")
#as.data.frame(lapply((pl.cam),hr.obs, as.numeric))
```


```{r}
#Qualitative combination
HRcuanmerging <- mn+obs
HRcualmerging=HRcuanmerging
HRcualmerging[HRcuanmerging!=0]=1

#Quantitative sum
HRcuanmerging <- mn+obs

# Grand total
mn.gts<- mn/sum(mn)
obs.gts<- obs/sum(obs)
HRgtmerging<- (mn.gts+obs.gts)/2
sum(HRgtmerging)

# Min-Max scaling
mn.mms<- (mn-min(mn))/(max(mn)-min(mn))
obs.mms<- (obs-min(obs))/(max(obs)-min(obs))
HRminmaxmerging<- (mn.mms+obs.mms)/2
```

## Bootstrap estimates of network parameters

We estimate the confidence intervals by resampling the five adjacency matrices
obtained from the different methods and computing their average, SE, and CIs.


### Cualitative matrix sum

```{r CIs_HRcualmerging}
#
# HRcualmerging = Cualitative matrix sum

require(bootstrap)
require(bipartite)
require(sampling)
#
# Bootstrap loop to create the boot matrices --------------

TIME <- Sys.time()
mat <- HRcualmerging
HRobs<- as.vector(networklevel(mat, 
    			  index = c("connectance",
    			  		    "NODF",
    						"weighted NODF",
    						"weighted connectance")))

# Resampling plan for bootstrapping n=99 adjacency matrices.
resamp <- 99                # Number of bootstrap samples required
mnmetrics_boot<- NULL

# The loop. PUT THIS IN A BOOT LOOP 
for (i in 1:resamp) {
    mat1 <- sample_frac(mat, 0.8, replace= T)
    mnmetrics<- as.vector(networklevel(mat1, 
    			  index = c("connectance",
    			  		    "NODF",
    						"weighted NODF",
    						"weighted connectance")))
    mnmetrics_boot<- rbind(mnmetrics_boot, mnmetrics)   
}
	colnames(mnmetrics_boot)<- c("connectance", "NODF", 
								 "weighted NODF",
								 "weighted connectance")

# Outputs bootstraped parameter differences values as dataframe. -------------
HRcualmerging_boot <- as.data.frame(colMeans(mnmetrics_boot))
rnd.se <- as.vector(apply(mnmetrics_boot,2,sd)/sqrt(resamp)) # SE randomized parameters
HRcualmerging_boot<- cbind(HRcualmerging_boot, rnd.se)
# ci1 <-  t(as.vector(apply(mnmetrics_boot,2,mean)-1.962*HRcualmerging_boot$rnd.se))
# ci2 <-  t(as.matrix(apply(mnmetrics_boot,2,mean)+1.962*HRcualmerging_boot$rnd.se))
ci1 <-  HRobs-(1.962*HRcualmerging_boot$rnd.se)
ci2 <-  HRobs+(1.962*HRcualmerging_boot$rnd.se)

HRcualmerging_boot<- cbind(HRcualmerging_boot, ci1, ci2)
colnames(HRcualmerging_boot)<- c("mean.boot", "SE", 
								 "CI_low",
								 "CI_high")
HRcualmerging_boot<- cbind(HRobs,HRcualmerging_boot)

#
# Output parameters
# meandf         # Bootstrap mean of parameter values
# SE             # Bootstrap SE for parameter values
# CI_low         # Bootstrap CI (lower) 
# CI_high        # Bootstrap CI (higher) 
HRcualmerging_boot

#
Sys.time() - TIME
#

```

### Cuantitative matrix sum

```{r CIs_HRcuanmerging}
#
# HRcuanmerging = Cuantitative matrix sum

require(bootstrap)
require(bipartite)
require(sampling)
#
# Bootstrap loop to create the boot matrices --------------

TIME <- Sys.time()
mat <- HRcuanmerging
HRobs<- as.vector(networklevel(mat, 
    			  index = c("connectance",
    			  		    "NODF",
    						"weighted NODF",
    						"weighted connectance")))

# Resampling plan for bootstrapping n=99 adjacency matrices.
resamp <- 99                # Number of bootstrap samples required
mnmetrics_boot<- NULL

# The loop. PUT THIS IN A BOOT LOOP 
for (i in 1:resamp) {
    mat1 <- sample_frac(mat, 0.8, replace= T)
    mnmetrics<- as.vector(networklevel(mat1, 
    								   index = c("connectance",
    									   		  "NODF",
    									   		  "weighted NODF",
    							 	   		  "weighted connectance")))
    mnmetrics_boot<- rbind(mnmetrics_boot, mnmetrics)   
}
	colnames(mnmetrics_boot)<- c("connectance", "NODF", 
								 "weighted NODF",
								 "weighted connectance")

# Outputs bootstraped parameter differences values as dataframe. -------------
HRcuanmerging_boot <- as.data.frame(colMeans(mnmetrics_boot))
rnd.se <- as.vector(apply(mnmetrics_boot,2,sd)/sqrt(resamp)) # SE randomized parameters
HRcuanmerging_boot<- cbind(HRcuanmerging_boot, rnd.se)
# ci1 <-  t(as.vector(apply(mnmetrics_boot,2,mean)-1.962*HRcuanmerging_boot$rnd.se))
# ci2 <-  t(as.matrix(apply(mnmetrics_boot,2,mean)+1.962*HRcuanmerging_boot$rnd.se))
ci1 <-  HRobs-(1.962*HRcuanmerging_boot$rnd.se)
ci2 <-  HRobs+(1.962*HRcuanmerging_boot$rnd.se)

HRcuanmerging_boot<- cbind(HRcuanmerging_boot, ci1, ci2)
colnames(HRcuanmerging_boot)<- c("mean.boot", "SE", 
								 "CI_low",
								 "CI_high")
HRcuanmerging_boot<- cbind(HRobs,HRcuanmerging_boot)

#
# Output parameters
# meandf         # Bootstrap mean of parameter values
# SE             # Bootstrap SE for parameter values
# CI_low         # Bootstrap CI (lower) 
# CI_high        # Bootstrap CI (higher) 
HRcuanmerging_boot

#
Sys.time() - TIME
#

```

### Grand Total Standardization

```{r CIs_HRgtmerging}
#
# HRgtmerging = Grand Total Standardization

require(bootstrap)
require(bipartite)
require(sampling)
#
# Bootstrap loop to create the boot matrices --------------

TIME <- Sys.time()
mat <- HRgtmerging
HRobs<- as.vector(networklevel(mat, 
    			  index = c("connectance",
    			  		    "NODF",
    						"weighted NODF",
    						"weighted connectance")))

# Resampling plan for bootsrapping n=99 adjacency matrices.
resamp <- 99                # Number of bootstrap samples required
mnmetrics_boot<- NULL

# The loop. PUT THIS IN A BOOT LOOP 
for (i in 1:resamp) {
    mat1 <- sample_frac(mat, 0.8, replace= T)
    mnmetrics<- as.vector(networklevel(mat1, 
    								   index = c("connectance",
    									   		 "NODF",
    									   		 "weighted NODF",
    							 	   		     "weighted connectance")))
    mnmetrics_boot<- rbind(mnmetrics_boot, mnmetrics)   
}
	colnames(mnmetrics_boot)<- c("connectance", "NODF", 
								 "weighted NODF",
								 "weighted connectance")

# Outputs bootstraped parameter differences values as dataframe. -------------
HRgtmerging_boot <- as.data.frame(colMeans(mnmetrics_boot))
rnd.se <- as.vector(apply(mnmetrics_boot,2,sd)/sqrt(resamp)) # SE randomized parameters
HRgtmerging_boot<- cbind(HRgtmerging_boot, rnd.se)
# ci1 <-  t(as.vector(apply(mnmetrics_boot,2,mean)-1.962*HRgtmerging_boot$rnd.se))
# ci2 <-  t(as.matrix(apply(mnmetrics_boot,2,mean)+1.962*HRgtmerging_boot$rnd.se))
ci1 <-  HRobs-(1.962*HRgtmerging_boot$rnd.se)
ci2 <-  HRobs+(1.962*HRgtmerging_boot$rnd.se)

HRgtmerging_boot<- cbind(HRgtmerging_boot, ci1, ci2)
colnames(HRgtmerging_boot)<- c("mean.boot", "SE", 
							   "CI_low", "CI_high")

HRgtmerging_boot<- cbind(HRobs,HRgtmerging_boot)

#
# Output parameters
# meandf         # Bootstrap mean of parameter values
# SE             # Bootstrap SE for parameter values
# CI_low         # Bootstrap CI (lower) 
# CI_high        # Bootstrap CI (higher) 
HRgtmerging_boot

#
Sys.time() - TIME
#


```

### Min-max scaled matrix

```{r CIs_HRminmaxmerging}
#
# HRminmaxmerging = Min-max scaled matrix

require(bootstrap)
require(bipartite)
require(sampling)
#
# Bootstrap loop to create the boot matrices --------------

TIME <- Sys.time()
mat <- HRminmaxmerging
HRobs<- as.vector(networklevel(mat, 
    			  index = c("connectance",
    			  		    "NODF",
    						"weighted NODF",
    						"weighted connectance")))

# Resampling plan for bootsrapping n=99 adjacency matrices.
resamp <- 99                # Number of bootstrap samples required
mnmetrics_boot<- NULL

# The loop. PUT THIS IN A BOOT LOOP 
for (i in 1:resamp) {
    mat1 <- sample_frac(mat, 0.8, replace= T)
    mnmetrics<- as.vector(networklevel(mat1, 
    								   index = c("connectance",
    									   		  "NODF",
    									   		  "weighted NODF",
    							 	   	  	      "weighted connectance")))
    mnmetrics_boot<- rbind(mnmetrics_boot, mnmetrics)   
}
	colnames(mnmetrics_boot)<- c("connectance", "NODF", 
								 "weighted NODF",
								 "weighted connectance")

# Outputs bootstraped parameter differences values as dataframe. -------------
HRminmaxmerging_boot <- as.data.frame(colMeans(mnmetrics_boot))
rnd.se <- as.vector(apply(mnmetrics_boot,2,sd)/sqrt(resamp)) # SE randomized parameters
HRminmaxmerging_boot<- cbind(HRminmaxmerging_boot, rnd.se)
# ci1 <-  t(as.vector(apply(mnmetrics_boot,2,mean)-1.962*HRminmaxmerging_boot$rnd.se))
# ci2 <-  t(as.matrix(apply(mnmetrics_boot,2,mean)+1.962*HRminmaxmerging_boot$rnd.se))
ci1 <-  HRobs-(1.962*HRminmaxmerging_boot$rnd.se)
ci2 <-  HRobs+(1.962*HRminmaxmerging_boot$rnd.se)

HRminmaxmerging_boot<- cbind(HRminmaxmerging_boot, ci1, ci2)
colnames(HRminmaxmerging_boot)<- c("mean.boot", "SE", 
								 "CI_low",
								 "CI_high")

HRminmaxmerging_boot<- cbind(HRobs,HRminmaxmerging_boot)

#
# Output parameters
# meandf         # Bootstrap mean of parameter values
# SE             # Bootstrap SE for parameter values
# CI_low         # Bootstrap CI (lower) 
# CI_high        # Bootstrap CI (higher) 
HRminmaxmerging_boot

#
Sys.time() - TIME
#

```

### MIST-NETTING MATRIX

```{r CIs_HR.mist-netting}
#

require(bootstrap)
require(bipartite)
require(sampling)
#
# Bootstrap loop to create the boot matrices --------------

TIME <- Sys.time()
mat <- mn
HRobs<- as.vector(networklevel(mat, 
    			  index = c("connectance",
    			  		    "NODF",
    						"weighted NODF",
    						"weighted connectance")))

# Resampling plan for bootsrapping n=99 adjacency matrices.
resamp <- 99                # Number of bootstrap samples required
mnmetrics_boot<- NULL

# The loop. PUT THIS IN A BOOT LOOP 
for (i in 1:resamp) {
    mat1 <- sample_frac(mat, 0.8, replace= T)
    mnmetrics<- as.vector(networklevel(mat1, 
    								   index = c("connectance",
    									   		  "NODF",
    									   		  "weighted NODF",
    							 	   	  	      "weighted connectance")))
    mnmetrics_boot<- rbind(mnmetrics_boot, mnmetrics)   
}
	colnames(mnmetrics_boot)<- c("connectance", "NODF", 
								 "weighted NODF",
								 "weighted connectance")

# Outputs bootstraped parameter differences values as dataframe. -------------
HRMNmerging_boot <- as.data.frame(colMeans(mnmetrics_boot))
rnd.se <- as.vector(apply(mnmetrics_boot,2,sd)/sqrt(resamp)) # SE randomized parameters
HRMNmerging_boot<- cbind(HRMNmerging_boot, rnd.se)
# ci1 <-  t(as.vector(apply(mnmetrics_boot,2,mean)-1.962*HRminmaxmerging_boot$rnd.se))
# ci2 <-  t(as.matrix(apply(mnmetrics_boot,2,mean)+1.962*HRminmaxmerging_boot$rnd.se))
ci1 <-  HRobs-(1.962*HRMNmerging_boot$rnd.se)
ci2 <-  HRobs+(1.962*HRMNmerging_boot$rnd.se)

HRMNmerging_boot<- cbind(HRMNmerging_boot, ci1, ci2)
colnames(HRMNmerging_boot)<- c("mean.boot", "SE", 
								 "CI_low",
								 "CI_high")

HRMNmerging_boot<- cbind(HRobs,HRMNmerging_boot)

#
# Output parameters
# meandf         # Bootstrap mean of parameter values
# SE             # Bootstrap SE for parameter values
# CI_low         # Bootstrap CI (lower) 
# CI_high        # Bootstrap CI (higher) 
HRMNmerging_boot

#
Sys.time() - TIME
#

```


### OBSERVATION MATRIX

```{r CIs_HR.observation}
#

require(bootstrap)
require(bipartite)
require(sampling)
#
# Bootstrap loop to create the boot matrices --------------

TIME <- Sys.time()
mat <- obs
HRobs<- as.vector(networklevel(mat, 
    			  index = c("connectance",
    			  		    "NODF",
    						"weighted NODF",
    						"weighted connectance")))

# Resampling plan for bootsrapping n=99 adjacency matrices.
resamp <- 99                # Number of bootstrap samples required
mnmetrics_boot<- NULL

# The loop. PUT THIS IN A BOOT LOOP 
for (i in 1:resamp) {
    mat1 <- sample_frac(mat, 0.8, replace= T)
    mnmetrics<- as.vector(networklevel(mat1, 
    								   index = c("connectance",
    									   		  "NODF",
    									   		  "weighted NODF",
    							 	   	  	      "weighted connectance")))
    mnmetrics_boot<- rbind(mnmetrics_boot, mnmetrics)   
}
	colnames(mnmetrics_boot)<- c("connectance", "NODF", 
								 "weighted NODF",
								 "weighted connectance")

# Outputs bootstraped parameter differences values as dataframe. -------------
HROBS_boot <- as.data.frame(colMeans(mnmetrics_boot))
rnd.se <- as.vector(apply(mnmetrics_boot,2,sd)/sqrt(resamp)) # SE randomized parameters
HROBS_boot<- cbind(HROBS_boot, rnd.se)
# ci1 <-  t(as.vector(apply(mnmetrics_boot,2,mean)-1.962*HRminmaxmerging_boot$rnd.se))
# ci2 <-  t(as.matrix(apply(mnmetrics_boot,2,mean)+1.962*HRminmaxmerging_boot$rnd.se))
ci1 <-  HRobs-(1.962*HROBS_boot$rnd.se)
ci2 <-  HRobs+(1.962*HROBS_boot$rnd.se)

HROBS_boot<- cbind(HROBS_boot, ci1, ci2)
colnames(HROBS_boot)<- c("mean.boot", "SE", 
								 "CI_low",
								 "CI_high")

HROBS_boot<- cbind(HRobs,HROBS_boot)

#
# Output parameters
# meandf         # Bootstrap mean of parameter values
# SE             # Bootstrap SE for parameter values
# CI_low         # Bootstrap CI (lower) 
# CI_high        # Bootstrap CI (higher) 
HROBS_boot
```