---
title: "Network indices confidence intervals - El Puntal"
author: "Pedro Jordano"
date: "11/17/2020"
output: html_document
---

## Matrices

Final matrix Names:

`bc = Original Barcoding matrix from Pistacia lentiscus`   
`cualmerging = Cualitative matrix sum`        
`cuanmerging = Cuantitative matrix sum`          
`gtmerging = Grand Total Standarization`         
`minmaxmerging = Min-max scaled matrix`            
`seffortmerging = Sampling effort merging by mean among standardized matrices.`       


```{r load-data}
# Load matrices

pl.bc<- read.csv("../Case_studies_datasets/pl_bc_clean.csv",header=T,sep=",",dec=".")
pl.cam<- read.csv("../Case_studies_datasets/pl_cam_clean.csv",header=T,sep=",",dec=".")

# Convert all interactions in numeric
bc <- as.data.frame(lapply(pl.bc, as.numeric))%>% 
	column_to_rownames(var="X")
# ??column_to_rownames
cam <- as.data.frame(lapply(pl.cam, as.numeric)) %>% 
	        column_to_rownames(var="X")

# Qualitative combination
PUNcuanmerging<- bc+cam
PUNcualmerging= PUNcuanmerging
PUNcualmerging[PUNcuanmerging!= 0]= 1

# Quantitative sum
PUNcuanmerging<- bc+cam

# Grand total
bct<- bc/sum(bc)
camt<- cam/sum(cam)
PUNgtmerging<- (bct+camt)/2

# Min-Max scaling
bct<- (bc-min(bc))/(max(bc)-min(bc))
camt<- (cam-min(cam))/(max(cam)-min(cam))
PUNminmaxmerging<- (bct+camt)/2

# Sampling effort standardization
bceff<- read.csv("../Case_studies_datasets/bc_sampling_effort.csv",header=T,sep=";",dec=".", comment.char = "#")
cameff<- read.csv("../Case_studies_datasets/cam_sampling_effort.csv",header=T, sep=",",dec=".")

# I need to convert all data in visits/h/plant
# First TIME TRANSFORMATION
bct1 <- bc[,,]/(12*bceff[,7]) #I  HAVE NUMBER OF DAYS THE SEED TRAY WAS IN THE PLANT - DIVIDE BY DAYS*12 =renders visits/h (12 hours of the day, not 24)
camt1 <- cam[,,]/(cameff[,3]/60) # I HAVE MINS THE CAMS WAS RECODING

# Then AREA TRANSFORMATION 
bceff$correc_fac <- bceff$plant_area/bceff$area_seed_trap # I CALCULATE A 
                             # CORRECTION FACTOR i.e. plant area/seed tray area
bct2 <- bct1[,,]*bceff[,8] #I multiply it by the feces found
camt2 <- camt1[,,]*2 # I could multiply CAM DATA by 2 since the camera has only 
                     # recorded half of the plant - SHOULD I? (Not included)

# Then merging
PUNseffortmerging<- (bct2+camt2)/2

```


### Cualitative matrix sum

```{r CIs_PUNcualmerging}
#
# PUNcualmerging = Cualitative matrix sum

require(bootstrap)
require(bipartite)
require(sampling)
#
# Bootstrap loop to create the boot matrices --------------

TIME <- Sys.time()
mat <- PUNcualmerging
PUNobs<- as.vector(networklevel(mat, 
    			  index = c("connectance",
    			  		    "NODF",
    						"weighted NODF",
    						"weighted connectance")))

# Resampling plan for bootstrapping n=99 adjacency matrices.
resamp <- 99                # Number of bootstrap samples required
mnmetrics_boot<- NULL

# The loop. PUT THIS IN A BOOT LOOP 
for (i in 1:resamp) {
    mat1 <- sample_frac(mat, 0.8, replace= T)
    mnmetrics<- as.vector(networklevel(mat1, 
    								   index = c("connectance",
    									   		  "NODF",
    									   		  "weighted NODF",
    								   		      "weighted connectance")))
    mnmetrics_boot<- rbind(mnmetrics_boot, mnmetrics)   
}
	colnames(mnmetrics_boot)<- c("connectance", "NODF", 
								 "weighted NODF",
								 "weighted connectance")

# Outputs bootstraped parameter differences values as dataframe. -------------
PUNcualmerging_boot <- as.data.frame(colMeans(mnmetrics_boot))
rnd.se <- as.vector(apply(mnmetrics_boot,2,sd)/sqrt(resamp)) # SE randomized parameters
PUNcualmerging_boot<- cbind(PUNcualmerging_boot, rnd.se)
# ci1 <-  t(as.vector(apply(mnmetrics_boot,2,mean)-1.962*PUNcualmerging_boot$rnd.se))
# ci2 <-  t(as.matrix(apply(mnmetrics_boot,2,mean)+1.962*PUNcualmerging_boot$rnd.se))
ci1 <-  PUNobs-(1.962*PUNcualmerging_boot$rnd.se)
ci2 <-  PUNobs+(1.962*PUNcualmerging_boot$rnd.se)

# PUNcualmerging_boot<- cbind(PUNcualmerging_boot, t(ci1), t(ci2))
PUNcualmerging_boot<- cbind(PUNcualmerging_boot, ci1, ci2)
colnames(PUNcualmerging_boot)<- c("mean.boot", "SE", 
								  "CI_low",
								  "CI_high")
PUNcualmerging_boot<- cbind(PUNobs,PUNcualmerging_boot)

#
# Output parameters
# meandf         # Bootstrap mean of parameter values
# SE             # Bootstrap SE for parameter values
# CI_low         # Bootstrap CI (lower) 
# CI_high        # Bootstrap CI (higher) 
PUNcualmerging_boot

#
Sys.time() - TIME
#

```

### Cuantitative matrix sum

```{r CIs_PUNcuanmerging}
#
# PUNcuanmerging = Cuantitative matrix sum

require(bootstrap)
require(bipartite)
require(sampling)
#
# Bootstrap loop to create the boot matrices --------------

TIME <- Sys.time()
mat <- PUNcuanmerging
PUNobs<- as.vector(networklevel(mat, 
    			  index = c("connectance",
    			  		    "NODF",
    						"weighted NODF",
    						"weighted connectance")))

# Resampling plan for bootstrapping n=99 adjacency matrices.
resamp <- 99                # Number of bootstrap samples required
mnmetrics_boot<- NULL

# The loop. PUT THIS IN A BOOT LOOP 
for (i in 1:resamp) {
    mat1 <- sample_frac(mat, 0.8, replace= T)
    mnmetrics<- as.vector(networklevel(mat1, 
    								   index = c("connectance",
    									   		 "NODF",
    									   		 "weighted NODF",
    							 	   		     "weighted connectance")))
    mnmetrics_boot<- rbind(mnmetrics_boot, mnmetrics)   
}
	colnames(mnmetrics_boot)<- c("connectance", "NODF", 
								 "weighted NODF",
								 "weighted connectance")

# Outputs bootstrapped parameter differences values as dataframe. -------------
PUNcuanmerging_boot <- as.data.frame(colMeans(mnmetrics_boot))
rnd.se <- as.vector(apply(mnmetrics_boot,2,sd)/sqrt(resamp)) # SE randomized parameters
PUNcuanmerging_boot<- cbind(PUNcuanmerging_boot, rnd.se)
# ci1 <-  t(as.vector(apply(mnmetrics_boot,2,mean)-1.962*PUNcuanmerging_boot$rnd.se))
# ci2 <-  t(as.matrix(apply(mnmetrics_boot,2,mean)+1.962*PUNcuanmerging_boot$rnd.se))
ci1 <-  PUNobs-(1.962*PUNcuanmerging_boot$rnd.se)
ci2 <-  PUNobs+(1.962*PUNcuanmerging_boot$rnd.se)

# PUNcuanmerging_boot<- cbind(PUNcuanmerging_boot, t(ci1), t(ci2))
PUNcuanmerging_boot<- cbind(PUNcuanmerging_boot, ci1, ci2)
colnames(PUNcuanmerging_boot)<- c("mean.boot", "SE", 
								 "CI_low",
								 "CI_high")
PUNcuanmerging_boot<- cbind(PUNobs,PUNcuanmerging_boot)

#
# Output parameters
# meandf         # Bootstrap mean of parameter values
# SE             # Bootstrap SE for parameter values
# CI_low         # Bootstrap CI (lower) 
# CI_high        # Bootstrap CI (higher) 
PUNcuanmerging_boot

#
Sys.time() - TIME
#
   
```

### Grand Total Standardization

```{r CIs_PUNgtmerging}
#
# PUNgtmerging = Grand Total Standardization

require(bootstrap)
require(bipartite)
require(sampling)
#
# Bootstrap loop to create the boot matrices --------------

TIME <- Sys.time()
mat <- PUNgtmerging
PUNobs<- as.vector(networklevel(mat, 
    			  index = c("connectance",
    			  		    "NODF",
    						"weighted NODF",
    						"weighted connectance")))

# Resampling plan for bootsrapping n=99 adjacency matrices.
resamp <- 99                # Number of bootstrap samples required
mnmetrics_boot<- NULL

# The loop. PUT THIS IN A BOOT LOOP 
for (i in 1:resamp) {
    mat1 <- sample_frac(mat, 0.8, replace= T)
    mnmetrics<- as.vector(networklevel(mat1, 
    								   index = c("connectance",
    									   		 "NODF",
    									   		 "weighted NODF",
    							 	   		     "weighted connectance")))
    mnmetrics_boot<- rbind(mnmetrics_boot, mnmetrics)   
}
	colnames(mnmetrics_boot)<- c("connectance", "NODF", 
								 "weighted NODF",
								 "weighted connectance")

# Outputs bootstraped parameter differences values as dataframe. -------------
PUNgtmerging_boot <- as.data.frame(colMeans(mnmetrics_boot))
rnd.se <- as.vector(apply(mnmetrics_boot,2,sd)/sqrt(resamp)) # SE randomized parameters
PUNgtmerging_boot<- cbind(PUNgtmerging_boot, rnd.se)
# ci1 <-  t(as.vector(apply(mnmetrics_boot,2,mean)-1.962*PUNgtmerging_boot$rnd.se))
# ci2 <-  t(as.matrix(apply(mnmetrics_boot,2,mean)+1.962*PUNgtmerging_boot$rnd.se))
ci1 <-  PUNobs-(1.962*PUNgtmerging_boot$rnd.se)
ci2 <-  PUNobs+(1.962*PUNgtmerging_boot$rnd.se)

# PUNgtmerging_boot<- cbind(PUNgtmerging_boot, t(ci1), t(ci2))
PUNgtmerging_boot<- cbind(PUNgtmerging_boot, ci1, ci2)
colnames(PUNgtmerging_boot)<- c("mean.boot", "SE", 
								 "CI_low",
								 "CI_high")

PUNgtmerging_boot<- cbind(PUNobs,PUNgtmerging_boot)

#
# Output parameters
# meandf         # Bootstrap mean of parameter values
# SE             # Bootstrap SE for parameter values
# CI_low         # Bootstrap CI (lower) 
# CI_high        # Bootstrap CI (higher) 
PUNgtmerging_boot

#
Sys.time() - TIME
#


```

### Min-max scaled matrix

```{r CIs_PUNminmaxmerging}
#
# PUNminmaxmerging = Min-max scaled matrix

require(bootstrap)
require(bipartite)
require(sampling)
#
# Bootstrap loop to create the boot matrices --------------

TIME <- Sys.time()
mat <- PUNminmaxmerging
PUNobs<- as.vector(networklevel(mat, 
    			  index = c("connectance",
    			  		    "NODF",
    						"weighted NODF",
    						"weighted connectance")))

# Resampling plan for bootsrapping n=99 adjacency matrices.
resamp <- 99                # Number of bootstrap samples required
mnmetrics_boot<- NULL

# The loop. PUT THIS IN A BOOT LOOP 
for (i in 1:resamp) {
    mat1 <- sample_frac(mat, 0.8, replace= T)
    mnmetrics<- as.vector(networklevel(mat1, 
    								   index = c("connectance",
    									   		  "NODF",
    									   		  "weighted NODF",
    							 	   	  	      "weighted connectance")))
    mnmetrics_boot<- rbind(mnmetrics_boot, mnmetrics)   
}
	colnames(mnmetrics_boot)<- c("connectance", "NODF", 
								 "weighted NODF",
								 "weighted connectance")

# Outputs bootstraped parameter differences values as dataframe. -------------
PUNminmaxmerging_boot <- as.data.frame(colMeans(mnmetrics_boot))
rnd.se <- as.vector(apply(mnmetrics_boot,2,sd)/sqrt(resamp)) # SE randomized parameters
PUNminmaxmerging_boot<- cbind(PUNminmaxmerging_boot, rnd.se)
# cci1 <-  t(as.vector(apply(mnmetrics_boot,2,mean)-1.962*PUNminmaxmerging_boot$rnd.se))
# cci2 <-  t(as.matrix(apply(mnmetrics_boot,2,mean)+1.962*PUNminmaxmerging_boot$rnd.se))
ci1 <-  PUNobs-(1.962*PUNminmaxmerging_boot$rnd.se)
ci2 <-  PUNobs+(1.962*PUNminmaxmerging_boot$rnd.se)

PUNminmaxmerging_boot<- cbind(PUNminmaxmerging_boot, ci1, ci2)
colnames(PUNminmaxmerging_boot)<- c("mean.boot", "SE", 
								 "CI_low",
								 "CI_high")

PUNminmaxmerging_boot<- cbind(PUNobs,PUNminmaxmerging_boot)

#
# Output parameters
# meandf         # Bootstrap mean of parameter values
# SE             # Bootstrap SE for parameter values
# CI_low         # Bootstrap CI (lower) 
# CI_high        # Bootstrap CI (higher) 
PUNminmaxmerging_boot

#
Sys.time() - TIME
#

```


### Sampling effort standadrdized matrix

Sampling effort merging by mean among standardized matrices.
 
```{r CIs_PUNseffortmerging}
#
# PUNseffortmerging = Min-max scaled matrix

require(bootstrap)
require(bipartite)
require(sampling)
#
# Bootstrap loop to create the boot matrices --------------

TIME <- Sys.time()
mat <- PUNseffortmerging
PUNobs<- as.vector(networklevel(mat, 
    			  index = c("connectance",
    			  		    "NODF",
    						"weighted NODF",
    						"weighted connectance")))

# Resampling plan for bootsrapping n=99 adjacency matrices.
resamp <- 99                # Number of bootstrap samples required
mnmetrics_boot<- NULL

# The loop. PUT THIS IN A BOOT LOOP 
for (i in 1:resamp) {
    mat1 <- sample_frac(mat, 0.8, replace= T)
    mnmetrics<- as.vector(networklevel(mat1, 
    								   index = c("connectance",
    									   		  "NODF",
    									   		  "weighted NODF",
    							 	   	  	      "weighted connectance")))
    mnmetrics_boot<- rbind(mnmetrics_boot, mnmetrics)   
}
	colnames(mnmetrics_boot)<- c("connectance", "NODF", 
								 "weighted NODF",
								 "weighted connectance")

# Outputs bootstraped parameter differences values as dataframe. -------------
PUNseffortmerging_boot <- as.data.frame(colMeans(mnmetrics_boot))
rnd.se <- as.vector(apply(mnmetrics_boot,2,sd)/sqrt(resamp)) # SE randomized parameters
PUNseffortmerging_boot<- cbind(PUNseffortmerging_boot, rnd.se)
# cci1 <-  t(as.vector(apply(mnmetrics_boot,2,mean)-1.962*PUNseffortmerging_boot$rnd.se))
# cci2 <-  t(as.matrix(apply(mnmetrics_boot,2,mean)+1.962*PUNseffortmerging_boot$rnd.se))
ci1 <-  PUNobs-(1.962*PUNseffortmerging_boot$rnd.se)
ci2 <-  PUNobs+(1.962*PUNseffortmerging_boot$rnd.se)

# PUNseffortmerging_boot<- cbind(PUNseffortmerging_boot, t(ci1), t(ci2))
PUNseffortmerging_boot<- cbind(PUNseffortmerging_boot, ci1, ci2)
colnames(PUNseffortmerging_boot)<- c("mean.boot", "SE", 
							 	     "CI_low",
								     "CI_high")

PUNseffortmerging_boot<- cbind(PUNobs,PUNseffortmerging_boot)

#
# Output parameters
# meandf         # Bootstrap mean of parameter values
# SE             # Bootstrap SE for parameter values
# CI_low         # Bootstrap CI (lower) 
# CI_high        # Bootstrap CI (higher) 
PUNseffortmerging_boot

#
Sys.time() - TIME
#

```



### DNA-Barcoding matrix
```{r CIs_Barcoding}
#
# 

require(bootstrap)
require(bipartite)
require(sampling)
#
# Bootstrap loop to create the boot matrices --------------

TIME <- Sys.time()
mat <- bc
PUNobs<- as.vector(networklevel(mat, 
    			  index = c("connectance",
    			  		    "NODF",
    						"weighted NODF",
    						"weighted connectance")))

# Resampling plan for bootsrapping n=99 adjacency matrices.
resamp <- 99                # Number of bootstrap samples required
mnmetrics_boot<- NULL

# The loop. PUT THIS IN A BOOT LOOP 
for (i in 1:resamp) {
    mat1 <- sample_frac(mat, 0.8, replace= T)
    mnmetrics<- as.vector(networklevel(mat1, 
    								   index = c("connectance",
    									   		  "NODF",
    									   		  "weighted NODF",
    							 	   	  	      "weighted connectance")))
    mnmetrics_boot<- rbind(mnmetrics_boot, mnmetrics)   
}
	colnames(mnmetrics_boot)<- c("connectance", "NODF", 
								 "weighted NODF",
								 "weighted connectance")

# Outputs bootstraped parameter differences values as dataframe. -------------
PUNBC_boot <- as.data.frame(colMeans(mnmetrics_boot))
rnd.se <- as.vector(apply(mnmetrics_boot,2,sd)/sqrt(resamp)) # SE randomized parameters
PUNBC_boot<- cbind(PUNBC_boot, rnd.se)
# cci1 <-  t(as.vector(apply(mnmetrics_boot,2,mean)-1.962*PUNminmaxmerging_boot$rnd.se))
# cci2 <-  t(as.matrix(apply(mnmetrics_boot,2,mean)+1.962*PUNminmaxmerging_boot$rnd.se))
ci1 <-  PUNobs-(1.962*PUNBC_boot$rnd.se)
ci2 <-  PUNobs+(1.962*PUNBC_boot$rnd.se)

PUNBC_boot<- cbind(PUNBC_boot, ci1, ci2)
colnames(PUNBC_boot)<- c("mean.boot", "SE", 
								 "CI_low",
								 "CI_high")

PUNBC_boot<- cbind(PUNobs,PUNBC_boot)

#
# Output parameters
# meandf         # Bootstrap mean of parameter values
# SE             # Bootstrap SE for parameter values
# CI_low         # Bootstrap CI (lower) 
# CI_high        # Bootstrap CI (higher) 
PUNBC_boot

```


### GO-PRO matrix
```{r CIs_GoPro}
#
# 

require(bootstrap)
require(bipartite)
require(sampling)
#
# Bootstrap loop to create the boot matrices --------------

TIME <- Sys.time()
mat <- cam
PUNobs<- as.vector(networklevel(mat, 
    			  index = c("connectance",
    			  		    "NODF",
    						"weighted NODF",
    						"weighted connectance")))
#In table 2, connectance value is calculated by dividing 91 pairwise interactions in goPro monitoring by 640 which is the nomber of possible pairwise interactions considering the matrix size of the merged matrix (in this case the same size than DNA-Barcoding matrix).

# Resampling plan for bootsrapping n=99 adjacency matrices.
resamp <- 99                # Number of bootstrap samples required
mnmetrics_boot<- NULL

# The loop. PUT THIS IN A BOOT LOOP 
for (i in 1:resamp) {
    mat1 <- sample_frac(mat, 0.8, replace= T)
    mnmetrics<- as.vector(networklevel(mat1, 
    								   index = c("connectance",
    									   		  "NODF",
    									   		  "weighted NODF",
    							 	   	  	      "weighted connectance")))
    mnmetrics_boot<- rbind(mnmetrics_boot, mnmetrics)   
}
	colnames(mnmetrics_boot)<- c("connectance", "NODF", 
								 "weighted NODF",
								 "weighted connectance")

# Outputs bootstraped parameter differences values as dataframe. -------------
PUNCAM_boot <- as.data.frame(colMeans(mnmetrics_boot))
rnd.se <- as.vector(apply(mnmetrics_boot,2,sd)/sqrt(resamp)) # SE randomized parameters
PUNCAM_boot<- cbind(PUNCAM_boot, rnd.se)
# cci1 <-  t(as.vector(apply(mnmetrics_boot,2,mean)-1.962*PUNminmaxmerging_boot$rnd.se))
# cci2 <-  t(as.matrix(apply(mnmetrics_boot,2,mean)+1.962*PUNminmaxmerging_boot$rnd.se))
ci1 <-  PUNobs-(1.962*PUNCAM_boot$rnd.se)
ci2 <-  PUNobs+(1.962*PUNCAM_boot$rnd.se)

PUNCAM_boot<- cbind(PUNCAM_boot, ci1, ci2)
colnames(PUNCAM_boot)<- c("mean.boot", "SE", 
								 "CI_low",
								 "CI_high")

PUNCAM_boot<- cbind(PUNobs,PUNCAM_boot)

#
# Output parameters
# meandf         # Bootstrap mean of parameter values
# SE             # Bootstrap SE for parameter values
# CI_low         # Bootstrap CI (lower) 
# CI_high        # Bootstrap CI (higher) 
PUNCAM_boot

```





------------------------------------------------------------------------------


```{r session, echo=T}
sessionInfo()

```


