---
title: "Hato Ratón correlations"
author: "Elena Quintero"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r library, warning=F, message=F}
library(tidyverse)
library(tibble)
library(RColorBrewer)
library(gridExtra)
library(ggplot2)
library(ggpubr)
```

Read data
```{r}
mn <- read.csv("../Case_studies_datasets/hr_mn_clean.csv",header=T,sep=",",dec=".")%>% 
	column_to_rownames(var="AASPECIES")
obs <- read.csv("../Case_studies_datasets/hr_obs_clean.csv",header=T,sep=",",dec=".")%>% 
	column_to_rownames(var="AASPECIES")
QC <- read.csv("../Data_merging/HR_QC.csv",header=T,sep=",",dec=".")%>% 
	column_to_rownames(var="X")
QS <- read.csv("../Data_merging/HR_QS.csv",header=T,sep=",",dec=".")%>% 
	column_to_rownames(var="X")
GTS <- read.csv("../Data_merging/HR_GTS.csv",header=T,sep=",",dec=".")%>% 
	column_to_rownames(var="X")
MMS <- read.csv("../Data_merging/HR_MMS.csv",header=T,sep=",",dec=".")%>% 
	column_to_rownames(var="X")
```

Apply long format to matrices:

```{r matrices to lists, include=F}
MN.long<- mn %>% 
  rownames_to_column(var="bird")%>%
  pivot_longer(-bird, names_to = "plant", values_to = "visits.mn")

OBS.long<- obs %>% 
  rownames_to_column(var="bird")%>%
  pivot_longer(-bird, names_to = "plant", values_to = "visits.obs")

QC.long<- QC %>% 
  rownames_to_column(var="bird")%>%
  pivot_longer(-bird, names_to = "plant", values_to = "visits.qc")

QS.long<- QS %>% 
  rownames_to_column(var="bird")%>%
  pivot_longer(-bird, names_to = "plant", values_to = "visits.qs")

GTS.long<- GTS %>% 
  rownames_to_column(var="bird")%>%
  pivot_longer(-bird, names_to = "plant", values_to = "visits.gts")

MMS.long<- MMS %>% 
  rownames_to_column(var="bird")%>%
  pivot_longer(-bird, names_to = "plant", values_to = "visits.mms")
```

Combine all merging methods together in a single dataset and calculate a Coefficient of Variation for 'Standardizing' methods

```{r, include=F}
all.long <- right_join(MN.long, OBS.long)%>%
  right_join(QC.long) %>%
  right_join(QS.long) %>% 
  right_join(GTS.long) %>%
  right_join(MMS.long) %>%
  dplyr::rowwise() %>% mutate(SD=sd(c(visits.gts,visits.mms)),
                              MEAN=mean(c(visits.gts,visits.mms)), CV=SD/MEAN)
```


## COMPARING BETWEEN QUANTITATIVE METHODS

```{r comparing between methods, warning=F, message=F, echo=F}
QS_GTS <- ggplot(data=all.long, aes(x=(visits.qs), y=(visits.gts)))+geom_point()+
  theme_bw()+xlab(" QS")+ylab(" GTS")+
  #geom_smooth(method=lm)+ 
  stat_cor(aes(label = ..r.label..), method="kendall")+
  geom_smooth(color="#0570b0", size=0.5)
QS_MMS <- ggplot(data=all.long, aes(x=(visits.qs), y=(visits.mms)))+geom_point()+
  theme_bw()+xlab(" QS")+ylab(" MMS")+
  #geom_smooth(method=lm)+ 
  stat_cor(aes(label = ..r.label..), method="kendall")+
  geom_smooth(color="#0570b0", size=0.5)
GTS_MMS <- ggplot(data=all.long, aes(x=(visits.gts), y=(visits.mms)))+geom_point()+
  theme_bw()+xlab(" GTS")+ylab(" MMS")+
  #geom_smooth(method=lm)+ 
  stat_cor(aes(label = ..r.label..), method="kendall")+
  geom_smooth(color="#0570b0", size=0.5)

grid.arrange(QS_GTS, QS_MMS, GTS_MMS, ncol=3)
```
outlier: *Pistacia lentiscus - Sylvia atricapilla*

Now with data log-transformed to control for 0 inflated matrices.
```{r comparing between methods 2, warning=F, message=F, echo=F}
QS_GTS <- ggplot(data=all.long, aes(x=log(visits.qs), y=log(visits.gts)))+geom_point()+
  theme_bw()+xlab("log QS")+ylab("log GTS")+
  #geom_smooth(method=lm)+ 
  stat_cor(aes(label = ..r.label..), method="kendall")+
  geom_smooth(color="#0570b0", size=0.5)
QS_MMS <- ggplot(data=all.long, aes(x=log(visits.qs), y=log(visits.mms)))+geom_point()+
  theme_bw()+xlab("log QS")+ylab("log MMS")+
  #geom_smooth(method=lm)+ 
  stat_cor(aes(label = ..r.label..), method="kendall")+
  geom_smooth(color="#0570b0", size=0.5)
GTS_MMS <- ggplot(data=all.long, aes(x=log(visits.gts), y=log(visits.mms)))+geom_point()+
  theme_bw()+xlab("log GTS")+ylab("log MMS")+
  #geom_smooth(method=lm)+ 
  stat_cor(aes(label = ..r.label..), method="kendall")+
  geom_smooth(color="#0570b0", size=0.5)

grid.arrange(QS_GTS, QS_MMS, GTS_MMS, ncol=3)
```

-   **All correlations showed in the graphs are kendall's tau (based in ranking for non-parametric data).**
-  All methods are relatively similar, since they are highly correlated. 
- Less correlated QS and MMS.

### METHODS CORRELATIONS
```{r cors}
pearson <- cor(x=all.long[,c(3,4,6:8)], method = "pearson")

kendall <- cor(x=all.long[,c(3,4,6:8)], method = "kendall")
```

Correlation plot:

```{r cors 2}
library(corrplot)
library(RColorBrewer)

col <- colorRampPalette(c("white", "white","white", "white","white", "white", "white",  "white", "white","#D9EF8B","#A6D96A","#66BD63"))

col2 <- colorRampPalette(c("white", "white","white", "white","white", "white", "white",  "white", "white","#edf8b1","#7fcdbb","#2c7fb8"))

corrplot(pearson, method = "color",type="upper",number.cex = .7,addCoef.col="black", tl.col = "black", col=col2(40), title="Pearson UP vs Kendall DOWN", tl.pos = "d",tl.srt = 45, cl.pos = "n")

corrplot(kendall, method = "color",type="lower",number.cex = .7,addCoef.col=T, tl.col = "black",  col=col2(40),diag = FALSE, tl.pos = "n", cl.pos = "n",add=T)
```

```{r, include=F}
# #with confidence intervals
# library(psych)
# cor.ci(x=all.long[,c(3,4,6:8)], method = "pearson")
# cor.ci(x=all.long[,c(3,4,6:8)], method = "kendall")
# f<-cor.plot.upperLowerCi(e)
```

### COMPARING METHODOLOGIES WITH MERGING METHODS

*Number of registers for each methodology*

Mist-Netting:
```{r, echo=F}
sum(mn)
```

Focal Observations:
```{r, echo=F}
sum(obs)
```

Here we observe the relationship between the 2 methodologies used in the field (logarithm converted):

```{r echo=F, warning=F, message=F, echo=F}
log_MN_OBS <- ggplot(data=all.long, aes(x=log(visits.mn), y=log(visits.obs)))+geom_point()+
  theme_bw()+xlab("log Mist-netting")+ylab("log Focal Observations")+
  #geom_smooth(color="violetred4", size=0.5)+ 
  geom_smooth(method=lm, color="#0570b0", size=0.5)+
  stat_cor(aes(label = ..r.label..), method="kendall")+  
  labs(title="Methodology comparision: Mist-netting vs. Observations")
log_MN_OBS 
```

All points standing on the x and y-axis are interactions undetected with one or the other method.

### How well each "merging method" values the different two methodologies

#### QUANTITATIVE SUM (QS)

Log transformed:

```{r  warning=F, message=F}
QS_mn_log <- ggplot(data=all.long, aes(x=log(visits.qs), y=log(visits.mn)))+geom_point()+
  theme_bw()+ylab("log Mist-netting")+xlab("log QS")+
  geom_smooth(color="#0570b0", size=0.5)+
  #geom_smooth(method=lm, color="violetred1", size=0.5)+ 
  stat_cor(aes(label = ..r.label..), method="kendall")

QS_obs_log <- ggplot(data=all.long, aes(x=log(visits.qs), y=log(visits.obs)))+geom_point()+
  theme_bw()+ylab("log Focal Obs")+xlab("log QS")+
  geom_smooth(color="#0570b0", size=0.5)+
  #geom_smooth(method=lm, color="violetred1", size=0.5)+ 
  stat_cor(aes(label = ..r.label..), method="kendall")

grid.arrange(QS_mn_log, QS_obs_log, ncol=2, top="How QS merging treats each metholodgy (logs)")
```

#### Grand Total Standarization (GTS)

Log transformed:

```{r echo=F,  warning=F, message=F}
gts_mn_log <- ggplot(data=all.long, aes(x=log(visits.gts), y=log(visits.mn)))+geom_point()+
  theme_bw()+ylab("log Mist-netting")+xlab("log gts")+
  geom_smooth(color="#0570b0", size=0.5)+
  #geom_smooth(method=lm, color="violetred1", size=0.5)+ 
  stat_cor(aes(label = ..r.label..), method="kendall")

gts_obs_log <- ggplot(data=all.long, aes(x=log(visits.gts), y=log(visits.obs)))+geom_point()+
  theme_bw()+ylab("log Focal Obs")+xlab("log gts")+
  geom_smooth(color="#0570b0", size=0.5)+
  #geom_smooth(method=lm, color="violetred1", size=0.5)+ 
  stat_cor(aes(label = ..r.label..), method="kendall")

grid.arrange(gts_mn_log, gts_obs_log, ncol=2, top="How GTS merging treats each metholodgy (logs)")
```

#### MIN-MAX SCALING (MMS)

Log transformed:

```{r echo=F,  warning=F, message=F}
mms_mn_log <- ggplot(data=all.long, aes(x=log(visits.mms), y=log(visits.mn)))+geom_point()+
  theme_bw()+ylab("log Mist-netting")+xlab("log mms")+
  geom_smooth(color="#0570b0", size=0.5)+
  #geom_smooth(method=lm, color="violetred1", size=0.5)+ 
  stat_cor(aes(label = ..r.label..), method="kendall")

mms_obs_log <- ggplot(data=all.long, aes(x=log(visits.mms), y=log(visits.obs)))+geom_point()+
  theme_bw()+ylab("log Focal Obs")+xlab("log mms")+
  geom_smooth(color="#0570b0", size=0.5)+
  #geom_smooth(method=lm, color="violetred1", size=0.5)+ 
  stat_cor(aes(label = ..r.label..), method="kendall")

grid.arrange(mms_mn_log, mms_obs_log, ncol=2, top="How MMS merging treats each metholodgy (logs)")
```

SUMMARY:

- Higher correlation (Kendall) of QS and GTS methods with the mist-netting data.